---
title: "Hands-on Excercise 1: Geospatial Data Wrangling with R"
editor: visual
---

# Overview

In this hands-on exercise, I learn how to import and wrangle geospatial data using appropriate R packages

## Getting Started

The code chunk below install and load [sf](https://r-spatial.github.io/sf/) and tidyverse packages into R environment

```{r}
pacman::p_load(sf,tidyverse)
  
```

## Importing Geospatial Data

### Importing polygon feature data in shape file format

```{r}
mpsz = st_read(dsn = "data/geospatial", 
               layer = "MP14_SUBZONE_WEB_PL")
```

### Importing polyline feature data in shapefile format

```{r}
cyclingpath = st_read(dsn = "data/geospatial", 
                         layer = "CyclingPath")
```

### Importing GIS data in kml format

```{r}
preschool = st_read("data/geospatial/pre-schools-location-kml.kml")
```

## Checking the content of a simple feature data frame

### Working with st_grometry()

```{r}
st_geometry(mpsz)
```

### Working with glimpse()

```{r}
glimpse(mpsz)
```

### working with head()

```{r}
head(mpsz, n=5)
```

## Plotting the Geospatial Data

```{r}
plot(mpsz)
```

```{r}
plot(st_geometry(mpsz))
```

```{r}
plot(mpsz["PLN_AREA_N"])
```

## Working with Projection

### Assigning EPSG code to a simple feature data frame

```{r}
st_crs(mpsz)
```

```{r}
mpsz3414 <- st_set_crs(mpsz,3414)
```

```{r}
st_crs(mpsz3414)
```

Transforming the projection of preschool from wgs84 to svy 21

```{r}
preschool3414 <- st_transform(preschool, crs =3414)
```

## Importing and Converting An Aspatial Data

### Importing the aspatial data

```{r}
listings <- read_csv("data/aspatial/listings.csv")
```

```{r}
list(listings)
```

### Creating a simple feature data frame from an aspatial data frame

```{r}
listings_sf <- st_as_sf(listings,
                        coords = c("longitude","latitude"),
                        crs = 4326) %>%
  st_transform(crs = 3414)
```

```{r}
glimpse(listings_sf)
```

## Geoprocessing with sf package

### Buffering

```{r}
buffer_cycling <- st_buffer(cyclingpath,
                            dist = 5,
                            nQuadSegs = 30)
```

```{r}
buffer_cycling$Area <- st_area(buffer_cycling)
```

```{r}
sum(buffer_cycling$Area)
```

### Point-in-polygon count

```{r}
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
```

```{r}
summary(mpsz3414$`PreSch Count`)
```

```{r}
top_n(mpsz3414,1,'PreSch Count')
```

```{r}
mpsz3414$Area <- mpsz3414 %>%
  st_area()
```

```{r}
mpsz3414 <- mpsz3414 %>%
  mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)
```

Exploratory Data Analysis (EDA)

```{r}
hist(mpsz3414$`PreSch Density`)
```

```{r}
ggplot(data=mpsz3414, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning sub-zones with a single pre-school, on the other hand, \nthere are two planning sub-zones with at least 20 pre-schools",
      x = "Pre-school density (per km sq)",
      y = "Frequency")
```

```{r}
ggplot(data = mpsz3414, mapping = aes(x = as.numeric(mpsz3414$`PreSch Density`),y = mpsz3414$`PreSch Count`)) + geom_point()+ labs(x="Pre-school density(per km sq)",
                                                                                                                                   y= "Pre-school count")
```
